Binary companions of red giant stars in \apogee\ \DR

### 2018-02-28 paper/1-catalog/twoface-catalog.tex
Words

\newcommand{\nprior}{536,870,912}
\newcommand{\nposterior}{256}
\newcommand{\nstars}{96,231}
\newcommand{\nvisits}{397,559}
\newcommand{\ncompanions}{xx}

@@

\begin{abstract}\noindent % trust me
% Context
Repeat radial-velocity measurements of stars can be used to identify stellar,
sub-stellar, and planetary-mass companions.
Even a very small number of observation epochs can be useful for detecting companions,
though such data can be difficult to use for characterizing
individual systems:
There can be multiple qualitatively different orbital solutions that fit the data.
For these reasons, we custom-built a Monte Carlo sampler (\thejoker)
that delivers reliable (and often highly
multi-modal) posterior samplings for companion orbital parameters given sparse
radial-velocity data.
% Aims
Here we
perform a search for secondary companions of over 100,000 red-giant
stars observed in the
\apogee\ survey (\DR) with $\geq 3$ spectroscopic epochs.
% Methods
We select stars with probable companions by making a cut on our posterior belief about
the amplitude of the stellar radial-velocity variation induced by the orbit \todo{do we?}.
% Results
We provide (1)~a catalog of \todo{\ncompanions} companions for which the stellar companion
properties can be confidently determined in a posterior sense,
(2)~a catalog of stars \todo{YYY} stars that likely
have companions, but for which (in most cases) more observations
would be needed to uniquely
determine the orbital properties, and (3)~posterior samplings for all orbital-companion
parameters for all \nstars\
stars in the parent sample.
We highlight interesting systems and show the characteristics of stars with
confidently determined companion properties.
\end{abstract}

@@

\subsection{Control sample}
\label{sec:control-sample}

\todo{describe how we make it}

@@

@@

We don't expect there to be a sharp transition in orbital parameters between
stars with companions and stars without companions: There is a continuum of
companion properties.
For example, the companion masses can be low mass, or at long periods, or at
high inclination, all of which will make the velocity semi-amplitude, $K$, small
for a given system.
From the radial velocity data alone, there is therefore no way to define a
simple cut to select a complete sample of stars with companions.
Nevertheless, it is possible to define a cut that selects stars with
high-confidence companions while similarly minimizing an estimated
false-positive rate.

\todo{False-positive estimate using the null control sample.}

We select a sample of stars that confidently have companions using percentiles
computed from the posterior samples in the velocity semi-amplitude parameter,
$K$.
\figurename~\ref{fig:} shows distributions of percentiles in $\ln K$ computed
for all stars with posterior samplings from \thejoker.
\todo{percentiles figure}

\todo{Examples of stars with no companions at various num. epochs}

\todo{Examples of stars that meet K cut at various num. epochs}

>>>>

\newcommand{\nprior}{536,870,912}
\newcommand{\nposterior}{256}
\newcommand{\nstars}{96,231}
\newcommand{\nvisits}{397,559}
\newcommand{\ncontrol}{16,384}
\newcommand{\nunimodal}{XX}
\newcommand{\nmultimodal}{XX}

@@

\begin{abstract}\noindent % trust me
% Context
Repeat radial-velocity measurements of stars can be used to identify stellar,
sub-stellar, and planetary-mass companions.
Even a very small number of observation epochs can be useful for detecting companions,
though such data can be difficult to use for characterizing
individual systems:
There can be multiple qualitatively different orbital solutions that fit the data.
For these reasons, we custom-built a Monte Carlo sampler (\thejoker)
that delivers reliable (and often highly
multi-modal) posterior samplings for companion orbital parameters given sparse
radial-velocity data.
% Aims
Here we
perform a search for secondary companions of over 100,000 red-giant
stars observed in the
\apogee\ survey (\DR) with $\geq 3$ spectroscopic epochs.
% Methods
We select stars with probable companions by making a cut on our posterior belief
about the amplitude of the stellar radial-velocity variation induced by the
orbit.
% Results
We provide (1)~a catalog of \nunimodal\ companions for which the stellar
companion properties can be confidently determined in a posterior sense, (2)~a
catalog of stars \nmultimodal\ stars that likely have companions, but for which
(in most cases) more observations would be needed to uniquely determine the
orbital properties, and (3)~posterior samplings for all orbital-companion
parameters for all \nstars\ stars in the parent sample.
We highlight interesting systems and show the characteristics of stars with
confidently determined companion properties.
\end{abstract}

@@

\subsection{Null control sample}
\label{sec:control-sample}

We construct a control sample of simulated data with no RV variability to assess
our false-positive rates in the selections below (see
\sectionname~\ref{sec:results}).
We randomly pick \ncontrol\ stars from the parent sample used in this work and
replace the visit velocity measurements with simulated data.
For each star $n$, we randomly sample an excess variance parameter value from
the prior, $s_n$.
For each visit $k$, we then sample a new velocity $v_{nk}$ by drawing from a
Gaussian with mean equal to the mean of the real data visit velocities,
$\bar{v}_n$, and variance equal to the sum of the visit variance (uncertainty),
$\sigma_{nk}$, and the excess variance,
\begin{equation}
    v_{nk} \sim \mathcal{N}(\bar{v}_n, \sigma_{nk}^2 + s_n^2) \quad .
\end{equation}
When we save the control data, we only store the visit uncertainty and
``forget'' the fact that the simulated data is generated with excess variance.

@@

We also run on the null control sample (see \sectionname~\ref{sec:control}) with
the same parameters.
\todo{summary of this run - do any stars get flagged as needs MCMC/etc?}

@@

We don't expect there to be a sharp transition in orbital parameters between
stars with companions and stars without companions: There is a continuum of
companion properties.
For example, the companion masses can be low mass, or at long periods, or at
high inclination, all of which will make the velocity semi-amplitude, $K$, small
for a given system.
From the radial velocity data alone, there is therefore no way to define a
simple cut to select a complete sample of stars with companions.
Nevertheless, it is possible to define a cut that selects stars with
high-confidence companions.

We select a sample of stars that confidently have companions using percentiles
computed from the posterior samples in the velocity semi-amplitude parameter,
$K$.
\figurename~\ref{fig:lnK-percentiles}, left panel, shows distributions of
percentiles in $\ln K$ computed for all stars with posterior samplings from
\thejoker.
\figurename~\ref{fig:lnK-percentiles}, right panel, shows the same for posterior
samplings for the null control sample.
We define a cut in the \todo{XX} percentile of the $\ln K$ samples such that
\todo{XX false-positives...} (vertical line in
\figurename~\ref{fig:lnK-percentiles}); \nmultimodal\ stars pass this cut.

% % Notebook: "TODO"
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\textwidth]{lnK-percentiles}
% \end{center}
% \caption{%
% TODO
% \label{fig:lnK-percentiles}
% }
% \end{figure}

\tablename~\ref{tbl:} contains a subset of the \apogee\ \DR\ \texttt{allStar}
catalog information for the stars that pass the above cut and therefore likely
have companions; For brevity below, we refer to this sample as the ``high-$K$''
stars, and the complimentary sample as the ``low-$K$'' stars.
\figurename~\ref{fig:highK} shows eight examples of high-$K$ stars, i.e. stars
that likely have companions, with different numbers of visits.
\figurename~\ref{fig:lowK} is the same for eight low-$K$ stars, i.e. stars that
either have no companions, low-mass companions, or companions at long periods or
high inclination.

The majority of the stars in the high-$K$ sample have poorly constrained orbital
parameters because the posterior pdfs are very multimodal.

% % Notebook: "TODO"
% Should also include period-eccentricity panels with samples
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\textwidth]{highK}
% \end{center}
% \caption{%
% TODO
% \label{fig:highK}
% }
% \end{figure}

% % Notebook: "TODO"
% \begin{figure}[h]
% \begin{center}
% \includegraphics[width=\textwidth]{lowK}
% \end{center}
% \caption{%
% TODO
% \label{fig:lowK}
% }
% \end{figure}
